Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.

### 2024-02-28 docs/_tex/index.tex
word and pdf included

\affiliation[1]{organization={Nantes University},,postcodesep={}}
\affiliation[2]{organization={Aveiro University},,postcodesep={}}
\affiliation[3]{organization={CNR ISMAR},,postcodesep={}}
\affiliation[4]{organization={JeSaisPas},,postcodesep={}}

@@

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

@@

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

@@

\includegraphics[width=26.67in,height=\textheight]{./Figures/Fig1_Map_Drone_Sites.png}

@@

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

@@

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

@@

\textsubscript{Source:
\href{https://SigOiry.github.io/Drone_Paper_2023/index.qmd.html}{Article
Notebook}}

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Convolutional Neural Network (CNN) model with 2 hidden
layers. The training pixels were categorized into 7 different classes,
representing the various habitats encountered at the different study
sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae,
Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were
used for training purposes because of their 8 mm spatial resolution
allowing to avoid spectral sub-pixel mixing and to accurately visually
differentiate various vegetation classes. More than 418,000 pixels at 8
mm resolution from the 3 training flights have been used to train the
model (Table 2). Twenty one variables were used by the model as
predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX
multispectral camera (ranging from 444 nm to 840 nm), the same 10
spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{cao2017}) and the NDVI
(Equation~\ref{eq-ndvi}). Standardisation of spectral bands is used to
eliminate the scaling differences between spectra and to limit the
effect of biomass on the shape of the spectra \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

>>>>

\affiliation[1]{organization={Institut des Substances et Organismes de
la Mer, ISOMer, Nantes Université, UR 2160, F-44000 Nantes,
France},,postcodesep={}}
\affiliation[2]{organization={CESAM -- Centre for Environmental and
Marine Studies, Department of Biology, University of Aveiro, Aveiro,
Portugal},,postcodesep={}}
\affiliation[3]{organization={Consiglio Nazionale delle Ricerche,
Istituto di Scienze Marine (CNR-ISMAR), 00133 Rome,
Italy},,postcodesep={}}
\affiliation[4]{organization={BRGM French Geological Survey, Cayenne
97300, French Guiana},,postcodesep={}}

@@

@@

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Fig1_Map_Drone_Sites.png}

@@

@@

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Convolutional Neural Network (CNN) model with 2 hidden
layers. The training pixels were categorized into 7 different classes,
representing the various habitats encountered at the different study
sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae,
Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were
used for training purposes because of their 8 mm spatial resolution
allowing to avoid spectral sub-pixel mixing and to accurately visually
differentiate various vegetation classes. More than 418,000 pixels at 8
mm resolution from the 3 training flights have been used to train the
model (Table 2). Twenty one variables were used by the model as
predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX
multispectral camera (ranging from 444 nm to 840 nm), the same 10
spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{Cao2017}) and the NDVI
(Equation~\ref{eq-ndvi}). Standardisation of spectral bands is used to
eliminate the scaling differences between spectra and to limit the
effect of biomass on the shape of the spectra \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

### 2024-02-28 docs/index.qmd
word and pdf included

---
title: Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.
author:
  - name: Simon Oiry
    orcid: 0000-0001-7161-5246
    corresponding: true
    email: oirysimon@gmail.com
    roles:
      - Investigation
      - Writing
      - Data acquisition
    affiliations:
      - Nantes University
  - name: Bede Ffinian Rowe Davies
    corresponding: false
    roles: 
      - Investigation
      - Data acquisition
      - Revision
    affiliations:
      - Nantes University
  - name: Pierre Gernez
    corresponding: false
    roles: 
      - Investigation
      - Data acquisition
      - Revision
    affiliations:
      - Nantes University
  - name: Ana I. Sousa
    corresponding: false
    roles:
      - Data acquisition
      - Revision
    affiliations:
      - Aveiro University
  - name: Philippe Rosa
    corresponding: false
    roles: 
      - Data acquisition
      - Revision
    affiliations:
      - Nantes University
  - name: Maria Laura Zoffoli
    corresponding: false
    roles: 
      - Data acquisition
      - Revision
    affiliations:
      - CNR ISMAR
  - name: Guillaume Brunier
    corresponding: false
    roles: 
      - Data acquisition
      - Revision
    affiliations:
      - JeSaisPas
  - name: Laurent Barillé
    corresponding: false
    roles: 
      - Investigation
      - Data acquisition
      - Revision
    affiliations:
      - Nantes University
keywords:
  - Drone
  - Remote Sensing
abstract: |
  Coastal areas host seagrass meadows, which offer crucial ecosystem services including erosion control and carbon sequestration. However, these areas are increasingly impacted by human activities, leading to seagrass decline and habitat fragmentation. In situ surveys, traditionally performed to monitor these ecosystems face limitations on temporal and spatial coverage, particularly in intertidal zones, prompting the use of satellite data within monitoring programs. Yet, satellite remote sensing struggles with spatial and spectral resolution, making it difficult to discriminate seagrass from other macrophytes in highly heterogenous meadows. To address these challenges, drone images acquired with multi-spectral sensors offer a promising solution. This study focuses on using drones acquisitions for mapping intertidal macrophytes, effectively discriminating between seagrass and green macroalgae. Ten drone flights were conducted at two different altitudes (12m and 120m) across diverse European habitats in France and Portugal. Low altitude flights were used to train a Deep Learning classifier based on Neural Networks to discrimintate among 5 intertidal vegetation classes. Drone mapping demonstrated an overall accuracy of 94% across all the sites and images, covering a total area of 467 000 m². The model exhibited an accuracy of 96.4% in identifying seagrass.
plain-language-summary: |
  This study demonstrates the efficacy of using drones equipped with multi-spectral sensors for mapping intertidal macrophytes, particularly seagrass, overcoming limitations of traditional in situ surveys and satellite remote sensing, with a Deep Learning classifier achieving 94% overall accuracy across diverse European habitats, and 96.4% accuracy specifically in identifying seagrass.
key-points:
  - An intertidal vegetation classifier has been built to be used across different study sites without losing classification accuracy. 
  - Neural Network made to accuratly discriminates intertidal seagrasses from green algae (chlorophyceae)
date: last-modified
number-sections: true
---

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Sites where the different drone flights were made. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon. Light green represent terrestrial areas whereas darkgreen represent intertidal areas. 
#| label: fig-map
#| fig-width: 9
#| fig-height: 9

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Convolutional Neural Network (CNN) model with 2 hidden layers. The training pixels were categorized into 7 different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were used for training purposes because of their 8 mm spatial resolution allowing to avoid spectral sub-pixel mixing and to accurately visually differentiate various vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights have been used to train the model (Table 2). Twenty one variables were used by the model as predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same 10 spectral bands standardized using a min/max transformation ([@eq-std] ; @cao2017) and the NDVI (@eq-ndvi). Standardisation of spectral bands is used to eliminate the scaling differences between spectra and to limit the effect of biomass on the shape of the spectra [@Douay2022 ; @Davies2023].

>>>>

---
title: Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery.
author:
  - name: Simon Oiry
    orcid: 0000-0001-7161-5246
    corresponding: true
    email: oirysimon@gmail.com
    roles:
      - Investigation
      - Writing
      - Data acquisition
    affiliations:
      - Institut des Substances et Organismes de la Mer, ISOMer, Nantes Université, UR 2160, F-44000 Nantes, France
  - name: Bede Ffinian Rowe Davies
    corresponding: false
    roles: 
      - Investigation
      - Data acquisition
      - Revision
    affiliations:
      - Institut des Substances et Organismes de la Mer, ISOMer, Nantes Université, UR 2160, F-44000 Nantes, France
  - name: Pierre Gernez
    corresponding: false
    roles: 
      - Investigation
      - Data acquisition
      - Revision
    affiliations:
      - Institut des Substances et Organismes de la Mer, ISOMer, Nantes Université, UR 2160, F-44000 Nantes, France
  - name: Ana I. Sousa
    corresponding: false
    roles:
      - Data acquisition
      - Revision
    affiliations:
      - CESAM – Centre for Environmental and Marine Studies, Department of Biology, University of Aveiro, Aveiro, Portugal
  - name: Philippe Rosa
    corresponding: false
    roles: 
      - Data acquisition
      - Revision
    affiliations:
      - Institut des Substances et Organismes de la Mer, ISOMer, Nantes Université, UR 2160, F-44000 Nantes, France
  - name: Maria Laura Zoffoli
    corresponding: false
    roles: 
      - Data acquisition
      - Revision
    affiliations:
      - Consiglio Nazionale delle Ricerche, Istituto di Scienze Marine (CNR-ISMAR), 00133 Rome, Italy
  - name: Guillaume Brunier
    corresponding: false
    roles: 
      - Data acquisition
      - Revision
    affiliations:
      - BRGM French Geological Survey, Cayenne 97300, French Guiana
  - name: Laurent Barillé
    corresponding: false
    roles: 
      - Investigation
      - Data acquisition
      - Revision
    affiliations:
      - Institut des Substances et Organismes de la Mer, ISOMer, Nantes Université, UR 2160, F-44000 Nantes, France
keywords:
  - Drone
  - Remote Sensing
abstract: |
  Coastal areas host seagrass meadows, which offer crucial ecosystem services including erosion control and carbon sequestration. However, these areas are increasingly impacted by human activities, leading to seagrass decline and habitat fragmentation. In situ surveys, traditionally performed to monitor these ecosystems face limitations on temporal and spatial coverage, particularly in intertidal zones, prompting the use of satellite data within monitoring programs. Yet, satellite remote sensing struggles with spatial and spectral resolution, making it difficult to discriminate seagrass from other macrophytes in highly heterogenous meadows. To address these challenges, drone images acquired with multi-spectral sensors offer a promising solution. This study focuses on using drones acquisitions for mapping intertidal macrophytes, effectively discriminating between seagrass and green macroalgae. Ten drone flights were conducted at two different altitudes (12m and 120m) across diverse European habitats in France and Portugal. Low altitude flights were used to train a Deep Learning classifier based on Neural Networks to discrimintate among 5 intertidal vegetation classes. Drone mapping demonstrated an overall accuracy of 94% across all the sites and images, covering a total area of 467 000 m². The model exhibited an accuracy of 96.4% in identifying seagrass.
plain-language-summary: |
  This study demonstrates the efficacy of using drones equipped with multi-spectral sensors for mapping intertidal macrophytes, particularly seagrass, overcoming limitations of traditional in situ surveys and satellite remote sensing, with a Deep Learning classifier achieving 94% overall accuracy across diverse European habitats, and 96.4% accuracy specifically in identifying seagrass.
key-points:
  - An intertidal vegetation classifier has been built to be used across different study sites without losing classification accuracy. 
  - Neural Network made to accuratly discriminates intertidal seagrasses from green algae (chlorophyceae)
date: last-modified
citation:
  container-title: Earth and Space Science
number-sections: true
notebook-links: false
---

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Sites where the different drone flights were made. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon. Light green represent terrestrial areas whereas darkgreen represent intertidal areas. 
#| label: fig-map
#| fig-width: 7
#| fig-height: 7
#| out-width: "100%"

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Convolutional Neural Network (CNN) model with 2 hidden layers. The training pixels were categorized into 7 different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were used for training purposes because of their 8 mm spatial resolution allowing to avoid spectral sub-pixel mixing and to accurately visually differentiate various vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights have been used to train the model (Table 2). Twenty one variables were used by the model as predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same 10 spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the NDVI (@eq-ndvi). Standardisation of spectral bands is used to eliminate the scaling differences between spectra and to limit the effect of biomass on the shape of the spectra [@Douay2022 ; @Davies2023].
